---
title: 'R Notebook sandbox: Assignment "Datasets" (10 points)'
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 4
    fig_caption: true
    number_sections: true
my-var: "monte"  # https://bookdown.org/yihui/rmarkdown/html-document.html
---



```{r,mychunk-common-include,message=FALSE}
library(devtools);
my.source = 'local';
local.path = "C:/_git_/WSU_STATS419_FALL2020/";
local.data.path = ""
source( paste0(local.path,"functions/libraries.R"), local=T );


```


# Matrix

Create the "rotate matrix" functions as described in lectures.  Apply to the example "myMatrix".

```{r,mychunk-matrix-include,message=FALSE}
source( paste0(local.path,"functions/functions-matrix.R"), local=T );


myMatrix = matrix ( c (
											1, 0, 2,
											0, 3, 0,
											4, 0, 5
											), nrow=3, byrow=T);


```


```{r,mychunk-matrix-demo}

transposeMatrix(myMatrix);
rotateMatrix90(myMatrix);  # clockwise ... 
rotateMatrix180(myMatrix);
rotateMatrix270(myMatrix);

```




# IRIS
Recreate the graphic for the IRIS Data Set using R.  Same titles, same scales, same colors.  See: https://en.wikipedia.org/wiki/Iris_flower_data_set#/media/File:Iris_dataset_scatterplot.svg

```{r,mychunk-iris-demo}

palette(value = c("red" , "green3","blue"))
plot(iris[1:4], col = iris$Species, pch=19, main = "Iris Data (red=setosa,green=versicolor,blue=virginica)")

```

Edgar Anderson introduce a multivariate data set called the Iris dataset consisting of 5 attributes for Iris Flowers and a sample of 50 of each species measurements. The measurements are Sepal Length, Sepal Width, Petal Length, and Petal Width for each flower.


# Personality

## Cleanup RAW
Import "personality-raw.txt" into R.  Remove the V00 column.  Create two new columns from the current column "date_test":  year and week. Stack Overflow may help:  https://stackoverflow.com/questions/22439540/how-to-get-week-numbers-from-dates ... Sort the new data frame by YEAR, WEEK so the newest tests are first ... The newest tests (e.g., 2020 or 2019) are at the top of the data frame.  Then remove duplicates using the unique function based on the column "md5_email".  Save the data frame in the same "pipe-delimited format" ( | is a pipe ) with the headers.  You will keep the new data frame as "personality-clean.txt" for future work (you will not upload it at this time).  In the homework, for this tasks, report how many records your raw dataset had and how many records your clean dataset has.


```{r,mychunk-clean-up-raw}
my_data = read.table(paste0(local.path,"datasets/personality/personality-raw.txt"), sep ="|", header = TRUE, dec =".")
my_data = my_data[,-3]
my_data$date_test = as.Date(sub(" .*", "", my_data$date_test), format = '%m/%d/%Y')
my_data$WEEK = format(as.Date(my_data$date_test), "%W")
my_data$YEAR = format(as.Date(my_data$date_test), "%Y")
my_data = my_data[order(my_data$YEAR,my_data$WEEK, decreasing = TRUE),]
my_data_clean = distinct(my_data, md5_email, .keep_all = TRUE)
#write.table(my_data_clean, file = "C:/Users/blue/Desktop/STAT419/personality-clean.txt", sep = "|", row.names = TRUE)

```

Raw record had 838 records and the clean set had 678 records.

# Variance and Z-scores 

Write functions for doSummary and sampleVariance and doMode ... test these functions in your homework on the "monte.shaffer@gmail.com" record from the clean dataset.  Report your findings.  For this "monte.shaffer@gmail.com" record, also create z-scores.  Plot(x,y) where x is the raw scores for "monte.shaffer@gmail.com" and y is the z-scores from those raw scores.  Include the plot in your assignment, and write 2 sentences describing what pattern you are seeing and why this pattern is present.

```{r,mychunk-monte-dataset}

monte = my_data_clean[my_data_clean$md5_email == 'b62c73cdaf59e0a13de495b84030734e',]
monte

```

## doMode

```{r,mychunk-doMode}

doMode = function(x) #from stackoverflow https://stackoverflow.com/questions/2547402/how-to-find-the-statistical-mode
{
  ux <- unique(x)
  tab <- tabulate(match(x, ux))
  ux[tab == max(tab)]
}

```

## Variance

```{r,mychunk-variance} 

doSampleVariance = function(x,method='two-pass')
{
  #NAIVE METHOD
  if(method=='naive')
  {
    return(sum((x-mean(x))^2)/(length(x)-1));
  }
  else
  {
    #two-pass algo
    sumsq = 0
    for (i in x)
    {
      sumsq = sumsq + ((i - mean(x))*(i-mean(x)))
    }
    return(sumsq / (length(x)-1));
  }
}

```

## doSummary
```{r,mychunk-doSummary}

doSummary = function(x) #worked on this with Nic 
{
  #length
  print("dimension")
  print(dim(x))
  #number of NAs
  print("number of NAs")
  print(sum(is.na(x)))
  #mean
  print("mean of each column")
  print(colMeans(x))
  #median
  print("median of each column")
  print(colMedians(data.matrix(x)))
  #mode
  print("mode of each column")
  print(sapply(x,doMode))
  #variance
  print("variance of each column")
  print(sapply(x,doSampleVariance))
  #sd
  print("sd of each column")
  print(sapply(x,sd))
}

doSummary(my_data_clean[,3:62])

```

## Z-Scores

```{r,mychunk-zscore} 

zscore = function(x,y) #calculate zscore
{
  #(x-mean)/sd
  mean = colMeans(x)
  sd = sapply(x,sd)
  vec = c()
  for (i in 1:ncol(y))
  {
    vec = c(vec,(y[,i]-data.frame(mean[i]))/data.frame(sd[i]))
  }
  return(vec);
}

#zscore taking the measurable columns 
xlist = as.list(monte[,3:62])
ylist = zscore(my_data_clean[,3:62],monte[,3:62])
zscoretable = do.call(rbind, Map(data.frame, x=xlist, y=ylist))
plot(zscoretable)


```

Just plotting this does not tell me much besides the fact that the professor tends to pick a majority of his mood scores above 4.0 so the bell curve peak would be around the 3.5-4.5 point region. 

# Will vs Denzel
```{r,mychunk-imdb-include,message=FALSE}
source( paste0(local.path,"functions/functions-imdb.R"), local=T );

```

Compare Will Smith and Denzel Washington. [See 03_n greater 1-v2.txt for the necessary functions and will-vs-denzel.txt for some sample code and in DROPBOX: \__student_access__\unit_01_exploratory_data_analysis\week_02\imdb-example ]  You will have to create a new variable $millions.2000 that converts each movie's $millions based on the $year of the movie, so all dollars are in the same time frame.  You will need inflation data from about 1980-2020 to make this work.


## Will Smith
```{r,mychunk-will, fig.cap = c("Will Smith scatterplot: IMDB(2020)", "Will Smith boxplot raw millions: IMDB(2020)")}
nmid = "nm0000226";
 	will = grabFilmsForPerson(nmid);
 	plot(will$movies.50[,c(1,6,7:10)]);
  	boxplot(will$movies.50$millions);
		widx =  which.max(will$movies.50$millions);
	will$movies.50[widx,];
		summary(will$movies.50$year);  # bad boys for life ... did data change?
		
```
## Denzel Washington
```{r,mychunk-denzel}
nmid = "nm0000243";
 	denzel = grabFilmsForPerson(nmid);
 	plot(denzel$movies.50[,c(1,6,7:10)]);
  	boxplot(denzel$movies.50$millions);
		didx =  which.max(denzel$movies.50$millions);
	denzel$movies.50[didx,];
		summary(denzel$movies.50$year);
		
```

## BoxPlot of Top-50 movies using Raw Dollars

```{r,mychunk-boxplot-raw}
par(mfrow=c(1,2));
	boxplot(will$movies.50$millions, main=will$name, ylim=c(0,360), ylab="Raw Millions" );
	boxplot(denzel$movies.50$millions, main=denzel$name, ylim=c(0,360), ylab="Raw Millions" );
	
	par(mfrow=c(1,1));
		
```

## Side-by-Side Comparisons
Build side-by-side box plots on several of the variables (including #6) to compare the two movie stars.  After each box plot, write 2+ sentence describing what you are seeing, and what conclusions you can logically make.  You will need to review what the box plot is showing with the box portion, the divider in the box, and the whiskers.

### Adjusted Dollars (2000)
### Total Votes  (Divide by 1,000,000)
### Average Ratings
### Year?  Minutes?
### Metacritic (NA values)


